---
Notebook title: Cluster analysis of Hoechst fluorescent intensity in E.coli
Output: html_document
Project: Hoechst dynamics and kinetics
Date : 20211122

---
The dataset includes single-cell fluorescent measurements of Hoechst intensity in E.coli strain NCMM3722 (wild type). Data were personally collected and processed. Under the microscope, apparently
there'a heterogeneity in Hoechst signals. The objective of this notebook is to classify this so-called "Phenotype" based solely on intensity. This cluster analysis will serve as an entry point to
further investigate the kinetics of Hoechst accumulation and degradation.
Firstly, a screening of clustering algorithms and number of clusters are performed. Then, clustering is performed with optimal algorithms and optimal number of clusters to classify the phenotypes.
The results are then saved in a csv file.
## Load packages
```{r}
library(tidyverse)
library(readxl)
library(plotly)
library(openxlsx)
library(gridExtra)
library(ggprism)
library(scales)
library(emojifont)
library(Cairo)
library(ggplot2)
library(ggnewscale)
library(stats)
library(sparklyr)
library(dplyr)
library(data.table)
library(clValid)
library(mclust)
library(tidymodels)
library(patchwork)
```
## Download data
```{r}
url <- "https://www.dropbox.com/s/ic053m6xixmag5t/HT_degradation_demo.csv?dl=1"
df_deg<- fread(url)
df_deg
```

## Cleaning up the dataframe. Remove unnecessary columns
```{r}
df1 <- subset( df_deg ,select = -c( IMAGE,
                                    INTENSITY,
                                    INTENSITY.ch2.mean_c,
                                    LOCATION,
                                    Final_OD600,
                                    Strain,
                                    Protocol,
                                    Date_and_growth_curve,
                                    Media,
                                    Exposure_time
                                    ))
# Remove time -1 (background measurements)
df2 <- df1[df1$Time_m != -1,]
df2
```
## Find optimal clustering algorithms with clValid package
```{r}
# normed_intensity_ch2 is fluorescent intensity after normalized with exposure time
# Choose algorithms need to be tested
clmethods <- c("hierarchical","kmeans", "diana", "fanny", "pam", "sota", "clara","model")
## Loop through all the concentrations to perform the tests
for (i in unique(df2$HT_Conc_uM)){
  data <- df2$normed_intensity_ch2[df2$HT_Conc_uM == i]
  pdf <- as.data.frame(as.numeric(data))
  val <- clValid(pdf, nClust = 2:4,clMethods = clmethods,
                 validation = "internal")

  print(paste("Currently working on HT concentration ",i,"uM"))
  print(optimalScores(val))
  plot(val,main = paste("HT concentration ",i,"uM"))
}
```
From the results, with all HT concentrations, hierarchical and k-means clustering algorithms are the best performers with optimal number of clusters is 2.
The following is phenotype classification with hclust and kmeans functions provided in stats package

```{r}
# Create vectors to store the results of clustering
K_pred <- c()
H_pred <- c()
# Loop through all the concentrations
for (i in unique(df2$HT_Conc_uM)){
  # filter for data
  data <- df2$normed_intensity_ch2[df2$HT_Conc_uM == i]
  # perform k-means clustering on data and set results as kclust
  kclust <- data %>% kmeans(centers = 2)
  # append k-means results to K_pred vector
  K_pred <- append(K_pred, kclust$cluster)
  # calculate dissimilarity matrix from data then create hclust object on dissimilarity
  # matrix with method = "average"
  hclust <- data %>%
                dist()%>%
                    hclust(method = "average")
  # use cutree to cut hclust object into 2 clusters
  cut_avg <- cutree(hclust, k = 2)
  # append classification results to H_pred
  H_pred <- append(H_pred, cut_avg)

}
```
## Plot histogram of HT intensity distributions with predicted phenotype classification
```{r}

# Join classification results to dataframe
df2$K_labels <- K_pred
df2$H_labels <- H_pred
# Create a vector of HT concentrations
htconc <- c(unique(df2$HT_Conc_uM))
for (i in seq_along(htconc)){
  data <- df2[df2$HT_Conc_uM == htconc[i],]
  p <- ggplot(data) +
          geom_histogram(aes(x= data$normed_intensity_ch2,
                        color = as.factor(data$K_labels)),
                        bins = 40,
                        show.legend = TRUE)+
          scale_y_log10(name = "Frequency")+
          scale_x_log10(name = "HT intensity")+
          scale_color_discrete(name = "Phenotype")+
          labs(title = paste("Classification with K-means for HT at ",htconc[i],"uM"))+
          theme(text=element_text(size=28))
  suppressWarnings(print(p))
}


for (i in seq_along(htconc)){
  data <- df2[df2$HT_Conc_uM == htconc[i],]
  q <- ggplot(data) +
          geom_histogram(aes(x= data$normed_intensity_ch2,
                                color = as.factor(data$H_labels)),
                                bins = 40,
                                show.legend = TRUE) +
          scale_y_log10(name = "Frequency")+
          scale_x_log10(name = "HT intensity")+
          scale_color_discrete(name = "Phenotype")+
          labs(title = paste("Classification with hierarchical clustering for HT at ",htconc[i],"uM"))+
          theme(text=element_text(size=28))
  suppressWarnings(print(q))
}
```
From the histograms, I can see that hierarchical clustering has better performance since there's only
1 phenotype for each given bin. In the case of k-means clustering, there's a mix of phenotypes
in some particular bins. Also from the histograms, I can name cluster 1 as "dark" and cluster 2 as "bright" at concentration 0.5, 1, 20 uM, and cluster 1 as "bright", cluster 2 as "dark" for
concentrations 3, 10, 15 uM.
Next, I will replace cluster indices with phenotype names
```{r}
# First, remove K_labels columns
df2 <- subset(df2,select= - K_labels)
# Replace 1 with "bright" and 2 with "dark" in H_labels columns at concentrations 3,10,15 uM
df2$H_labels[(df2$H_labels == 1) & (df2$HT_Conc_uM == 3 |
                                    df2$HT_Conc_uM == 10|
                                    df2$HT_Conc_uM == 15)] <- "bright"
df2$H_labels[df2$H_labels == 2 & (  df2$HT_Conc_uM == 3 |
                                    df2$HT_Conc_uM == 10|
                                    df2$HT_Conc_uM == 15)] <- "dark"
# Replace 1 with "dark" and 2 with "bright" in H_labels columns at concentrations 0.5,1,20 uM
df2$H_labels[(df2$H_labels == 2) & (df2$HT_Conc_uM == 0.5 |
                                    df2$HT_Conc_uM == 1   |
                                    df2$HT_Conc_uM == 20)] <- "bright"
df2$H_labels[df2$H_labels == 1 & (df2$HT_Conc_uM == 0.5 |
                                  df2$HT_Conc_uM == 1   |
                                  df2$HT_Conc_uM == 20)]   <- "dark"
# Change column name from H_labels to Classified_phenotype
colnames(df2)[7] <- "Classified_phenotype"
# Check data frame
df2
```
## Save dataframe in csv file
```{r}
write.csv(df2,"D:/Dropbox_dailescifi/Dropbox/RProject/ClusterHT.csv", row.names = FALSE)
```

