---
title: "Off rate of HT dissociation"
output: html_notebook
---
This notebook is used to calculate koff of HT dissociation based on dissociation assays. 
```{r}
library(plyr)
library(tidyverse)
library(dplyr)
library(data.table)
library(ggplot2)
library(stats)
library(sparklyr)
library(readxl)
library(plotly)
library(openxlsx)
library(gridExtra)
library(ggprism)
library(scales)
library(emojifont)
library(Cairo)
library(ggnewscale)
library(stats)
library(tidymodels)
library(patchwork)
library(formula.tools)
library(data.table)
```

```{r}
#Links to csv files
url1 <- "https://www.dropbox.com/s/ic053m6xixmag5t/HT_degradation_demo.csv?dl=1"
url2 <- "https://www.dropbox.com/s/di568r38yhduavw/HT_accumulation_demo.csv?dl=1"
url3 <- "https://www.dropbox.com/s/np77bgeuxh4sgsh/20210617_H342_accumulation_degradation_tabular_timelapse.xlsx?dl=1"
url4 <- "https://www.dropbox.com/s/132wq65kq1m7sv5/20211127_H342_accumulation_degradation_snapshot_from_old_data_to_tab.csv?dl=1"
url5 <- "https://www.dropbox.com/s/hn53irhedktloyg/20211208_H33342_degradation_timelapse_transfer_from_old.csv?dl=1"
# Read and parse to dataframe
df_deg <- fread(url1) # Data for degradation
df_acc <- fread(url2)
df_acct <- fread(url4)
df_degt <-fread(url5)
# Data for accumulation
# Check available columns
colnames(df_deg)
colnames(df_acc)
colnames(df_degt)
colnames(df_acct)


```
```{r}
df1 <- subset(df_acc ,select = c( Genotype,
                                  HT_Conc_uM,
                                  Repeat,
                                  Time_m,
                                  normed_intensity_ch2
))
df2 <- subset(df_deg, select = c( Genotype,
                                  HT_Conc_uM,
                                  Repeat,
                                  Time_m,
                                  normed_intensity_ch2 ))
df3 <- subset(df_degt , select =c( NAME,
                                   Genotype,
                                   HT_Conc_uM,
                                   Repeat,
                                   Time_m,
                                   normed_intensity_ch2,
                                   Phenotype))

# Remove time -1 (background measurements)
df_acc <- df_acc[order(df_acc$HT_Conc_uM,df_acc$Time_m),]
df_deg <- df_deg[order(df_deg$HT_Conc_uM,df_deg$Time_m),]
df_degt <- df_degt[order(df_degt$HT_Conc_uM,df_degt$Time_m),]
df_acct <- df_acct[order(df_acct$HT_Conc_uM,df_acct$Time_m),]
df_acc.WT <- df1[((df1$Time_m >= 0) &
  (df1$Genotype == "WT")&
  (df1$Repeat == 1)),]
df_acc.WT <- df_acc.WT[order(df_acc.WT$HT_Conc_uM,df_acc.WT$Time_m),]
df_deg.WT <- df2[((df2$Time_m >= 0) &
  (df2$Genotype == "WT") &
  (df2$Repeat == 1)),]
df_deg.WT <- df_deg.WT[order(df_deg.WT$HT_Conc_uM,df_deg.WT$Time_m),]
df_acct.WT <- df4[((df4$Time_m >= 0) &
  (df4$Genotype == "WT")&
  (df4$Repeat == 1)),]
df_acct.WT <- df_acct.WT[order(df_acct.WT$HT_Conc_uM,df_acct.WT$Time_m),]
df_degt.WT <- df3[((df3$Time_m >= 0) &
  (df3$Repeat == 1)),]

```
Calculate mean and standard deviation for each concentration at each time point
```{r}
df_deg.WT$"ln_int" <- log(df_deg.WT$normed_intensity_ch2)
df_deg.WT$"rev_int" <-1/df_deg.WT$normed_intensity_ch2
df_deg.WT.agg <- df_deg.WT  %>% group_by( HT_Conc_uM , Time_m ) %>%
        summarise(avg_int = mean(normed_intensity_ch2) ,
                  std_int = sd(normed_intensity_ch2),
                  avg_lnint = mean(ln_int),
                  std_lnint = sd(ln_int),
                  avg_1int = mean(rev_int),
                  std_1int = sd(rev_int))%>%
        mutate( minht    = avg_int - std_int,
                maxht    = avg_int + std_int,
                minlnht = avg_lnint - std_lnint,
                maxlnht = avg_lnint + std_lnint,
                min1int = avg_1int - std_1int,
                max1int = avg_1int + std_1int)

df_degt.WT$"ln_int" <- log(df_degt.WT$normed_intensity_ch2)
df_degt.WT$"rev_int" <-1/df_degt.WT$normed_intensity_ch2
df_degt.WT.agg <- df_degt.WT  %>% group_by( Phenotype, HT_Conc_uM , Time_m ) %>%
        summarise(avg_int = mean(normed_intensity_ch2) ,
                  std_int = sd(normed_intensity_ch2),
                  avg_lnint = mean(ln_int),
                  std_lnint = sd(ln_int),
                  avg_1int = mean(rev_int),
                  std_1int = sd(rev_int))%>%
        mutate( minht    = avg_int - std_int,
                maxht    = avg_int + std_int,
                minlnht = avg_lnint - std_lnint,
                maxlnht = avg_lnint + std_lnint,
                min1int = avg_1int - std_1int,
                max1int = avg_1int + std_1int)

```
Plot average of HT intensity as a function of time
```{r}
# Function for point plot 
pointplot <-function(df,x,y,ymin,ymax,color,shape,title,xtitle,ytitle){
  p<-ggplot (df, aes(x, y)) +
        ggnewscale :: new_scale_colour() +
        geom_point(aes(color = as.factor(color),
                       shape = shape),
                   size = 3,
                   stroke = 1.5) +
        geom_line(aes(color = as.factor(color),
                      shape = shape),
                  size = 1,
                  stroke = 1) +
        geom_errorbar(aes(x = x , 
                        y = y , 
                        color = as.factor(color) ,
                        ymin = ymin ,
                        ymax = ymax ),
                    width = 3,
                    size = 1 ) +
        scale_color_prism(palette = "prism_dark") +
        labs(title = title ) +
        scale_x_continuous(name = xtitle,
                           guide = 'prism_offset_minor' ,
                           position = "bottom",
                           limits = c(-2.5,62.5)
                           ) +
        scale_y_continuous(name = ytitle,
                           guide = "prism_offset_minor"
                           ) +
        theme_prism(palette = "prism_dark" ,base_size = 14)
  return(p)
}
```
```{r}
p1 <- pointplot(df = df_deg.WT.agg,
                x = df_deg.WT.agg$Time_m,
                y = df_deg.WT.agg$avg_int,
                ymin = df_deg.WT.agg$minht,
                ymax = df_deg.WT.agg$maxht,
                xtitle = 'Time(m)',
                ytitle = 'Intensity(A.U.)',
                title = 'HT degradation with WT',
                color = df_deg.WT.agg$HT_Conc_uM,
                shape = NULL)
p1
```
```{r}
p2 <- pointplot(df = df_deg.WT.agg,
                x = df_deg.WT.agg$Time_m,
                y = df_deg.WT.agg$avg_lnint,
                ymin = df_deg.WT.agg$minlnht,
                ymax = df_deg.WT.agg$maxlnht,
                xtitle = 'Time(m)',
                ytitle = 'LN Intensity(A.U.)',
                title = 'HT degradation with WT',
                color = df_deg.WT.agg$HT_Conc_uM,
                shape = NULL)
p2
```
```{r}
p3 <- pointplot(df = df_deg.WT.agg,
                x = df_deg.WT.agg$Time_m,
                y = df_deg.WT.agg$avg_1int,
                ymin = df_deg.WT.agg$min1int,
                ymax = df_deg.WT.agg$max1int,
                xtitle = 'Time(m)',
                ytitle = '1/HT Intensity(A.U.)',
                title = 'HT degradation with WT',
                color = df_deg.WT.agg$HT_Conc_uM,
                shape = NULL)
p3
```
k off of HT degradation kinetics at population level
```{r}
lreg1 <- function(df){
  reg<-c()
  r<-c()
  fitval<-c()
  summary<-c()
  ht<-c(list(unique(df$HT_Conc_uM)))
  for (i in unique(df$HT_Conc_uM)){
    data<-df[df$HT_Conc_uM == i,]
    reg<-append(reg,coef(lm(data$ln_int~data$Time_m))[1][1])
    sum<-summary(lm(data$ln_int~data$Time_m))
    r<-append(r,sum$r.squared)
    print(summary(lm(data$ln_int~data$Time_m)))
  }
  reg_df<-data.frame(ht,reg,r)
  colnames(reg_df)<-c("HT_conc","coefficients","r2")
  print(reg_df)
  return (reg_df)
}
```

```{r}
lreg2 <- function(df){
  reg<-c()
  r<-c()
  fitval<-c()
  summary<-c()
  ht<-c(list(unique(df$HT_Conc_uM)))
  for (i in unique(df$HT_Conc_uM)){
    data<-df[df$HT_Conc_uM == i,]
    reg<-append(reg,coef(lm(data$rev_int~data$Time_m))[1][1])
    sum<-summary(lm(data$rev_int~data$Time_m))
    r<-append(r,sum$r.squared)
    print(summary(lm(data$rev_int~data$Time_m)))
  }
  reg_df<-data.frame(ht,reg,r)
  colnames(reg_df)<-c("HT_conc","coefficients","r2")
  print(reg_df)
  return (reg_df)
}
```


```{r}
pop_reg1 <- lreg1(df_deg.WT[df_deg.WT$Time_m <=30,])
pop_reg2 <- lreg2 (df_deg.WT[df_deg.WT$Time_m <=30,])
```
At 1uM HT, 1st order kinetics fits better than 2nd order. Hence koff for 1uM HT
at population level is 3.36(A.U/min)

Since I already know that there's heterogeneity in HT signal at population level, 
plus high HT cells exhibit a different kinetics compared to low HT cells. It is 
indeed necessary to observe kinetics of HT degradation at single cell level with 
high HT and low HT cells analyzed separately.
```{r}
lineplot <-function(df,x,y,color,shape,title,xtitle,ytitle){
  p<-ggplot (df, aes(x, y)) +
        ggnewscale :: new_scale_colour() +
        geom_point(aes(color = as.factor(color),
                       shape = as.factor(shape)),
                   size = 3,
                   stroke = 1.5) +
        geom_line(aes(color = as.factor(color),
                      shape = as.factor(shape)),
                  size = 1,
                  stroke = 1) +
        labs(title = title ) +
        scale_x_continuous(name = xtitle,
                           guide = 'prism_offset_minor' ,
                           position = "bottom",
                           limits = c(-2.5,62.5)
                           ) +
        scale_y_continuous(name = ytitle,
                           guide = "prism_offset_minor"
                           ) +
        theme_prism(palette = "prism_dark" ,base_size = 14)
  return(p)
}
```

```{r}
unique(df_degt.WT$HT_Conc_uM)
```
```{r}
datat1 <- df_degt.WT %>% filter(df_degt.WT$HT_Conc_uM == 1 & df_degt.WT$Phenotype == 'dark')
p2 <- lineplot(df = datat1,
               x = datat1$Time_m,
               y = datat1$normed_intensity_ch2,
               color = datat1$NAME,
               shape = datat1$Phenotype,
               xtitle = "Time(m)",
               ytitle = "Intensity(A.U.)",
               title = "1uM HT degradation at single cell level")
p2
```
```{r}
p3 <- lineplot(df = datat1,
               x = datat1$Time_m,
               y = datat1$ln_int,
               color = datat1$NAME,
               shape = datat1$Phenotype,
               xtitle = "Time(m)",
               ytitle = "LN Intensity(A.U.)",
               title = "Ln 1uM HT degradation at single cell level")
p3
```
```{r}
p4 <- lineplot(df = datat1,
               x = datat1$Time_m,
               y = datat1$rev_int,
               color = datat1$NAME,
               shape = datat1$Phenotype,
               xtitle = "Time(m)",
               ytitle = "1/HT",
               title = "1/HT of 1uM HT degradation at single cell level")
p4
```
Linear regression 
```{r}
#Function for linear regression 

reg1 <- lreg1(datat1[datat1$Time_m <=30 ,])
```

```{r}

reg2 <- lreg2(datat1[datat1$Time_m <=30 ,])

```
From linear regression results, I can see that 1st order kinetic fits better
so k off is gradient of line ln(HT intensity) against t
Therefore k off for 1 uM HT degradation is 3.83(A.U/min)
This is similar to k off of 1uM HT degradation at population level (3.36)

Repeat the process for 10uM HT
```{r}
datat10 <- df_degt.WT %>% filter(df_degt.WT$HT_Conc_uM == 10 & df_degt.WT$Phenotype == 'dark')
p5 <- lineplot(df = datat10,
               x = datat10$Time_m,
               y = datat10$normed_intensity_ch2,
               color = datat10$NAME,
               shape = datat10$Phenotype,
               xtitle = "Time(m)",
               ytitle = "Intensity(A.U.)",
               title = "10uM HT degradation at single cell level")
p5
```
```{r}
p6 <- lineplot(df = datat10,
               x = datat10$Time_m,
               y = datat10$ln_int,
               color = datat10$NAME,
               shape = datat10$Phenotype,
               xtitle = "Time(m)",
               ytitle = "Intensity(A.U.)",
               title = "LN 10uM HT degradation at single cell level")
p6
```
```{r}
p7 <- lineplot(df = datat10,
               x = datat10$Time_m,
               y = datat10$rev_int,
               color = datat10$NAME,
               shape = datat10$Phenotype,
               xtitle = "Time(m)",
               ytitle = "Intensity(A.U.)",
               title = "1/HT of 10uM HT degradation at single cell level")
p7
```
Linear regression 
```{r}
reg10 <- lreg1(datat10)
reg20 <- lreg2(datat10)
```
Once again, 1st order kinetics fits better. Hence k off of 10uM HT degradation 
is 5.299 (A.U./min). This is also similar to k off of 10uM HT at population level
(5.4(A.U./min)). 




